---
title: "King County House prices Analysis by Sayari Ghosh"
author: "Sayari Ghosh"
date: "October 21, 2017"
output: github_document
always_allow_html: yes
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

This report explores the King county house prices dataset which is available on 
Kaggle website. In this analysis we will try to find out on which factors is the 
house price dependend on.  

```{r}
data = read.csv('kc_house_data.csv')
```

```{r}
head(data)
```

```{r}
dim(data)
```
Our dataset consists of 21 variables, with almost 21613 observations.

```{r}
str(data)
```

```{r}
summary(data)
```

```{r}
names(data)
```


The metatdata description of the variables is not provided in Kaggle. Although 
some of them are easy to understand there are a few which is ambigious.  
The dataset contains the following 21 variables: 

id : The id of the house           
date : The date when this information was taken.(dates between May 2014 
       and May 2015.)         
price : Price of the house       
bedrooms : No of bedrooms     
bathrooms : No of bathrooms     
sqft_living : Living area in square feet   
sqft_lot : Lot area in square feet   
floors : No of floors       
waterfront : House has a waterfront or not
view : Views of the houses         
condition : House condition ranging from 1 to 5    
grade : House grade ranging from 1 to 13        
sqft_above: Living area excluding the basement
sqft_basement : Basement area
yr_built : The year the house was build
yr_renovated : The year the house was renovated  
zipcode : Zipcode of the house
lat : Latitude 
long : Longitude    
sqft_living15 : The average house square footage of the 15 closest houses
sqft_lot15 : The average lot square footage of the 15 closest houses

The columns id,date are redundant in this analysis so it is better to remove
them.

```{r}
kc_data <- subset(data, , -c(id,date))
head(kc_data)
```

Preparing the dataset

Adding and modifying columns to be used later.

```{r}
kc_data$price_cat <- cut(kc_data$price,c(0,200000,500000,800000,1500000,7700000)
                     ,labels=c("Cheap", "Moderate", "High", "Expensive","Max"))

kc_data$yr_builtR[((kc_data$yr_built>=1900)&(kc_data$yr_built<=1925))]<-"1900-1925"
kc_data$yr_builtR[((kc_data$yr_built>=1926)&(kc_data$yr_built<=1950))]<-"1926-1950"
kc_data$yr_builtR[((kc_data$yr_built>=1951)&(kc_data$yr_built<=1975))]<-"1951-1975"
kc_data$yr_builtR[((kc_data$yr_built>=1976)&(kc_data$yr_built<=2000))]<-"1976-2000"
kc_data$yr_builtR[((kc_data$yr_built>=2001)&(kc_data$yr_built<=2015))]<-"2001-2015"
```

Creating a categorical variable for price and assigning them to "cheap","moderate","high","expensive","max". 
Creating categorical variable for the year_built column.
Created 5 categories for this. 

Converting the varibles waterfront,condition,floors,grade,bedrooms into factor
variables.

```{r}
kc_data$waterfront <- as.factor(kc_data$waterfront)
kc_data$condition <- as.factor(kc_data$condition)
kc_data$floors <- as.factor(kc_data$floors)
kc_data$grade <- as.factor(kc_data$grade)
kc_data$bedrooms <- as.factor(kc_data$bedrooms)
str(kc_data)
```

```{r}
head(kc_data)
```

Libraries used: 

```{r}
library(ggplot2)
library(ggthemes)
library(corrplot)
library(GGally)
library(magrittr)
library(dplyr)
library(zipcode)
library(plotly)
library(RColorBrewer)
library(lattice)
library(leaflet)
library(reshape2)
library(reshape)
library(scatterpie)
library(ROCR)
library(htmlwidgets)
```

# Univariate Plots Section 

```{r}
p1 <- ggplot(aes(x=bedrooms),data=kc_data) +
 geom_bar(fill="blue",alpha=.5)  + xlim(0,10) + theme_calc()+ scale_x_discrete("bedrooms") + 
 labs(x="Bedrooms", y="Count", title="Number of houses with different bedrooms")
ggplotly(p1)
```

Out of the houses in King county a major portion of them are of 3 bedrooms which 
is ideal for small families.  

```{r}
p2 <- ggplot(aes(x=yr_built),data=kc_data) +
  geom_histogram(col="red", fill="green", alpha = .2,binwidth = 1) + 
  theme_hc() +
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  labs(x="Year built", y="Count", title="Number of houses built in each year") 
ggplotly(p2)
```

There is a gradual increase in the number of houses built across the years 
with the maximum number of houses built in 2014. 

```{r}
p3 <- ggplot(aes(x=condition),data=kc_data) +
  geom_bar(fill="blue",alpha=.5)  + 
  xlim(0,8) + theme_calc()+  scale_x_discrete("condition") +
  labs(x="Condition", y="Count", title="Number of houses with different conditions") 
ggplotly(p3)
```

Most of the houses have a condition of 3 in King County. 

```{r}
p4 <- ggplot(aes(x=bathrooms),data=kc_data) +
  geom_bar(fill="blue",alpha=.5)  + 
  xlim(0,8) + theme_calc()+  scale_x_discrete("bathrooms") +
  labs(x="bathrooms", y="Count", title="Number of houses with different bathrooms") 
ggplotly(p4)
```

The most common bathrooms are 1 full bath and 2 full bath and 1 powder room in 
the houses in king county. 

```{r}
p5 <- ggplot(kc_data, aes(x=grade)) +
  geom_bar(fill='blue',stat="count",alpha=.5)  + theme_calc()+ scale_x_discrete("bedrooms") + 
  labs(x="grade", y="count", title="Number of houses with different grades")
ggplotly(p5)
```

The most common grade is 7 in King county with grades 1,2,3,13 being nonexistent

```{r}
summary(kc_data$sqft_living)
```

The median sqft_living is 1910 sq feet with the mean being 2080 sq feet.

```{r}
summary(kc_data$sqft_lot)
```

The median sqft_lot is 7618 sq feet with the mean being 15107 sq feet. The max 
lot size is quite huge compared to the the other houses and may be considered 
as an outlier.

```{r}
summary(kc_data$price)
```

The median price of the houses in king county is around 45000. 

```{r}
ggplot(kc_data,aes(x=price))+
  geom_density(fill='red') + xlim(0,4e+06) + 
  labs(x="price", y="density", title="Density distribution of price") 
```

The density distribution gives a plot that is right skewed. 

```{r}
ggplot(kc_data,aes(x=log(price)))+
  geom_density(fill='red') +
  labs(x="price", y="density", title="Density distribution of price in log scale") 
```

When the price is change to log of price the distribution changes to normal 
distrbution from a skewed distribution. 

```{r}
ggplot(kc_data, aes(x=sqft_living)) +
  geom_density(fill='blue') + xlim(0,10000) +
  labs(x="sqft_living", y="density", title="Density distribution of sqft living")
```

Density distribution for sqft_living of the houses and it is right skewed.

```{r}
zip_count <- kc_data %>%
  group_by(zipcode) %>%
  summarise(count = n()) 
data(zipcode)
zip_count <- merge(zip_count, zipcode, by.x='zipcode', by.y='zip')

head(zip_count[order(-zip_count$count),c(1,2,3)],5)
```


These 5 zipcodes have the highest number of houses and the top 3 cities are 
Seattle, Maple valley and Redmond. 

# Univariate Analysis

### What is the structure of your dataset?

There are 21613 houses in the dataset with 21 features. The variables condition,
grade, and view, waterfront are ordered factor variables. 
From the analysis done it seems that the most common condition is 3. The median 
price of the houses is 45000,the most common bedroom is 3,the most common grade 
is 7. 
As it is not mentioned in the metatdata which of the values of the factor 
variables is better it has to be found out later in bivariate plots. 

### What is/are the main feature(s) of interest in your dataset? 

The main features in the data set are price and sqft_living. I'd like to 
determine which features are best for predicting the price of a houses and which 
features have a strong correlation with price. I suspect sqft_living,sqft_lot,
bedrooms and grade can help to build a predictive model to price of houses.

### What other features in the dataset do you think will help support your \
### investigation into your feature(s) of interest?

I would like to explore condition, sqft_above, yr_build,yr_renovated and zipcode 
features of the dataset and find insights from the data. 

### Did you create any new variables from existing variables in the dataset?

Yes I created price_cat,yr_buildR to have caegorical variables for conducting 
furthur plots.

### Of the features you investigated, were there any unusual distributions? \
###Did you perform any operations on the data to tidy, adjust, or change the 
### form of the data? If so, why did you do this?

I will convert waterfront,condition,floors,grade into factor variables for 
bivariate and multivariate analysis. 

# Bivariate Plots Section
 
```{r, fig.width=13,fig.height=6}
qqmath(~sqft_above | price_cat, kc_data) 
```

By this Q-Q plots for each price category we see the variable sqft_above has a
right skewed distribution.

```{r,fig.width=10,fig.height=4}
kc_c <- cor(kc_data[, "price"], kc_data[ , sapply(kc_data, is.numeric)]) 
corrplot(kc_c, method="number")
```


In this correlation matrix of the price with other variables, the most dominant 
correlations price has is with sqft_living(0.7), grade(0.67), sqft_above(0.61), sqft_living15(0.59) and bathrooms(0.53). This was different from what I had 
expected. I was hoping that bedrooms,zip_code,condition and sqft_lot should have 
a higher correlation with price. Zipcode has a negative correlation(-0.05) which
is unusual. The average sqft living of 15 nearby houses (sqft_living15) is quite 
relevant as in neighbourhood which has higher sqft living size, most of the 
houses will be of similar size and also price. 

So now I will focus on these variables with higher correlation. 

Now looking into the plots for those variables with higher correlation value.

```{r}
ggplot(aes(x=sqft_living,y=price),data=kc_data) + geom_point(color="pink") +
  stat_smooth(method="lm",se=F,alpha=0.6,size=1) + 
  labs(title="correlation of price vs sqft living",x="Price in dollars", y="Sqft living")
options(scipen=999)
```

Sqft_living has the highest correlation with price which is obvious and in the 
price vs sqft_living we see a linear relationship between the two. The 
regression line shows that the relationship is linear. 

```{r}
ggpairs(data=kc_data, columns=c(1,11),
    mapping = aes(color = "red"),
    axisLabels="show") + labs(title="correlation of price vs sqft above") 
```

This plots made using ggpairs gives a lot of information at one time. Similarly 
sqft_above has more or less a linear with price. The density function of 
sqft_above is also right skewed. This correlation is the third largest among 
the variables selected. 

```{r,fig.width=10}
ggplot(aes(x=as.factor(bathrooms),y=price),data=kc_data) + geom_boxplot(col="green") +
  labs(x="bathrooms", title="price vs bathrooms")
```

As bathrooms have a strong correlation with price we plotted this to further 
understand the relationship. As the umber of bathroom increase the spread 
increase of the boxplot indicating that the price difference is increasing. 
Also as the number of bathrooms increase the first quantile value price also 
goes up giving a positive correlation with price.

```{r}
p7 <- ggplot(kc_data, aes(x=as.factor(grade),y=price)) +
  geom_boxplot(col="blue") +
  labs(x="grade", title="price vs grades")
ggplotly(p7)
```

Similarly this plot shows a gradual positive correlation of price with 
increasing grade.Hence we can figure that houses with grade 13 is the best as it
is the costliest. There are a lot of outliers in the grades ranging from 7 to 11

```{r}
meanprice_bygrade <- kc_data %>%
  filter(price<=6000000) %>%
  group_by(grade) %>% 
 summarize(avgPrice = mean(price),count = n()) 
meanprice_bygrade
```

The above is also proved by this summary table. 

```{r}
meanprice_byfloors <- kc_data %>%
  filter(price<=6000000) %>%
  group_by(floors) %>% 
 summarize(avgPrice = mean(price))

ggplot(aes(y=avgPrice,x=floors),data=meanprice_byfloors) + geom_bar(stat="identity") +
  labs(y="avgprice", x="floors", title="Average price for each floor") 
```

The floor that has the highest average price is 2.5 followed by 3.5.

```{r}
p8 <- ggplot(kc_data, aes(group=grade,x=grade,y=sqft_above)) +
  geom_boxplot(col="deep pink") + 
  labs(x="grade", title="sqft above vs grades")
ggplotly(p8)
```

Now looking into the correlation between grades and sqft above, I see there is a 
positive correlation.

```{r}
ggplot(kc_data, aes(x=as.factor(view),y=price)) +
  geom_boxplot(col="deep pink") + 
  labs(x="grade", title="sqft above vs grades") 
```

From this plot I figured that view 4 is the best among the four. 

```{r, fig.width=8,fig.height=8}
ggplot(aes(group=bedrooms,x=bedrooms,y=price),data=data) +
  geom_boxplot(col="dark green")  + xlim(0,11) + 
  ggtitle("Price ranging for bedrooms") 
```

There are a lot of outliers in this boxplot for bedrooms 3,4,5,6 which indicate
that price and bedrooms are not that correlated as I had assumed before. 

```{r}
ggplot(aes(x=sqft_lot,y=price),data=kc_data) + geom_point(color="deep pink") +
  stat_smooth(method="lm",se=F,alpha=0.6,size=1.5) + 
  labs(title="relation between house prices vs sqft lot")
```

Relationship of price vs sqft lot is clearly not linear and hence sqft lot 
clearly does not affect price of houses in king county.

```{r}
ggplot(kc_data,aes(x=price, fill=waterfront , colour = waterfront))+
  geom_density(alpha=.2) +
  xlim(0,3e+06) +
  labs(x="price", y="density", title="Density distribution of price with respect to waterfront   present or not") 
```

In this density plot the distribution of the price is studied with respect to
waterfront being present or not. In the case where waterfront is not present the 
prices are lower with the spread being less compared to the normal distribution 
with waterfront where the price range is larger. 

```{r}
basement <- kc_data %>%
mutate(basement_present = ifelse(sqft_basement == 0, "Not present", "Present")) %>%
group_by(basement_present) %>%
  filter(price<=6000000) %>%
  summarise(total = n(),avg_price = mean(price)) 
```

```{r}
p5 <- ggplot(data = basement,aes(x=total,y=avg_price,color=basement_present)) +
  geom_point() + xlim(5000,14000) + 
  labs(x="sqft_living", y="density", title="Density distribution of sqft living")
ggplotly(p5)
```

In the neighbourhood, the number of houses having basement is less than the
number of housing not having basement. By comparing the average prices we see 
that it is higher for the houses with that have basement. It means most people 
live in houses that does not have basement as the prices is comparitively 
much higher. 

```{r, fig.width=10,fig.height=7}
spineplot(y=kc_data$price_cat,x=as.factor(kc_data$yr_builtR),
          xlab="year category", ylab="price category",main="Price Distribution over years")
```

In the time range of 1900-1925 houses were more expensive compared to 1926-2000. 
In 2000 there is again increase in the price range.

```{r, fig.width=10,fig.height=7}
spineplot(y=as.factor(kc_data$floors),x=as.factor(kc_data$yr_builtR),
          xlab="year category", ylab="floors",main="floors of house Distribution over years")

```

In the early 1900s(1900-1950) mostly houses were built with 1,1.5 floors. But,
gradually through the years in the 2000s this has changed and mostly houses are 
built with 2 floors and 3.5 floors is also on the high.

```{r, fig.width=10,fig.height=7}
spineplot(y=as.factor(kc_data$condition),x=as.factor(kc_data$yr_builtR),
          xlab="year category", ylab="condition",main="house condition distribution over the years")
```
By plotting condition vs the year category I found that most of the houses built 
in he 2000s are of condition 3 compared to the previous years where other
condition houses were also built. I can assume that condition 3 signifies 
a mid condition that can be afforable by the middle working class. 

#Bivariate Analysis 

###Talk about some of the relationships you observed in this part of the \
###investigation. How did the feature(s) of interest vary with other features in \
###the dataset?

Yes there were some relationships of interest. 
# Houses built in 2000s are mostly of condition 3. 
# Houses built in 2000s are mostly of 2 and 3.5 floors compared to 1900-1950 
which were 1 and 1.5 floors 

The feature of interest was price and it has a positive correlation with sqft_living,grade,sqft_above. 
Contrary to my belief, number of bedrooms, sqft-lot and zipcode does not have
a strong correlation with price. 

### Did you observe any interesting relationships between the other features \
### (not the main feature(s) of interest)?

In the early 1900s(1900-1950) mostly houses were built with 1,1.5 floors. But,
gradually through the years in the 2000s this has changed and mostly houses are 
built with 2 floors and 3.5 floors is also on the high.

### What was the strongest relationship you found?
The strongest relationship is of price vs sqft_living as it has a positive 
correlation of 0.7

#Multivariate Plots

```{r, fig.width=10,fig.height=5}
ggplot(kc_data, aes(sqft_living, fill = as.factor(floors))) +
  geom_density(alpha=.5) + xlim(0,9000) + 
  labs(x="sqft_living", y="density", title="Density distribution of sqft_living for each floor")
```

In this plot I tried to density distribution of sqft living for each floor and 
found that similar to price the height for 2.5 floor is maximum and variation of 
price is less as compared to the other floor factors.

```{r}
ggplot(kc_data,aes(x=price, fill=condition , colour = condition)) +
  geom_density(alpha=.3) + xlim(0,2.5e+06) +
  labs(x="price", y="density", title="Density distribution of price in for each condition")
```

In this distribution the price is compared for each condition value. For the 
condition 2 the variation of price is less compared to the the other conditions. 

```{r}
kc_data_ren <- kc_data %>%
  filter( yr_renovated>=2000) %>%
  group_by(yr_renovated) %>%
  summarise(meanpriceRenovated = mean(price))

kc_data_blt <- kc_data %>%
  filter( yr_built>=2000) %>%
  group_by(yr_built) %>%
  summarise(meanpriceBuilt = mean(price))
 
houses <- merge.data.frame(kc_data_blt,kc_data_ren,by.x = "yr_built", by.y = "yr_renovated")

for(i in 1:nrow(houses))
{ houses$no[i] = i}  

houses<-melt(houses,id=c("no","yr_built"))
houses_s <- houses[order(houses$yr_built),]
```

```{r}
ggplot(houses_s,aes(y=value,x=yr_built,fill=variable))+
     geom_bar(stat="identity",width=.5,position="dodge") +
  labs(x="year", y="price", title="MeanPrice for yrbuilt vs yrRenovated")
```


The mean prices of the houses that were built in 2000 to 2012 was way lower 
that those renovated in that period. But there is a gradual shift and the mean
price built has increased after 2012.

```{r}
p11 <- ggplot(aes(y=sqft_living,x=sqft_living15),data=kc_data) +
 geom_point(aes(col=as.factor(grade)),alpha=.7) +
  labs(x="sqft_living15", y="sqft_living", title="Sqft_living15 vs sqft_living for each grade")
ggplotly(p11)
```

There is a strong correlation between sqft_living and sqft_living15. The grades 
from 1 to 6 has usually smaller values for sqft_living and sqft_living15 and it 
gradually increases for grades 7 to 13. 

```{r}
p13 <- ggplot(kc_data,aes(x=sqft_living,y=price,col=yr_built,group=factor(yr_built)))+
  geom_point(alpha=.5)+ xlim(0,10000) +
  scale_color_gradientn(colors = terrain.colors(10)) +
  labs(x="sqft_living", y="price", title="Sqft_living15 vs price for each year category")
ggplotly(p13)
```

For the same sqft_living houses built in 1900 has higher price than the ones 
built in 2000s. This is very interesting observation. But houses made in 
larger sqft_living has increased in the 2000s.

```{r, fig.width=10,fig.height=10}
p14 <- ggplot(data=subset(kc_data,price<=6000000), aes(y=price,x=sqft_above)) +
  geom_line(aes(color=condition)) + facet_wrap( ~ condition, ncol=1)  +
  labs(x="sqft_above", y="price", title="Sqft_above vs price for each year condition")
ggplotly(p14)
``` 

This plot helps to understand that houses with condition 1,2 has both less range 
of price and sqft above compared to the other 3. 

```{r}
ggplot(kc_data,aes(factor(bedrooms),factor(floors), col=factor(grade)))+
  geom_jitter(alpha=0.5, size=2)+
  scale_colour_manual(values = c('yellow','pink','red','orange',"black","white",'green',"deep pink","violet","brown","grey",'blue')) +
  labs(title="Bedrooms vs. grade",x="Bedrooms",y="floors")
```

This plot of bedrooms and floors has horizontal strips. The majority of houses with 3 floors have grade 8. For floors 2, the majority is 8 and 9. For floors 1 and 1.5 the most common grade is 6 and 7. 

```{r}
kc_data$view <- as.factor(kc_data$view)
p19 <- ggplot(kc_data, aes(sqft_living, sqft_lot,color=floors)) +
  geom_point(size=2, aes(shape=view)) +  xlim(0,10000) + ylim(0,1300000) +
  ggtitle("sqft-living vs sqft_lot")
p19
```

The plots shows that lot size is quite high compared to the sqft living. the majority of houses are with 1 and 2 floors and between 2400 to 7000 sqft_living. Most of the houses are of view 0. 

```{r}
zipbycondition <- kc_data %>%
  filter(price<=6000000) %>%
  group_by(zipcode,condition) %>% 
 summarize(avgPrice = mean(price),count = n()) 

zipbycondition_m <- cast(zipbycondition, zipcode~condition)
zipbycondition_m[is.na(zipbycondition_m)] <- 0
zipbycondition_m <- lapply(zipbycondition_m, as.numeric)

myvars <- c("lat", "long", "zipcode")
kc_new <- kc_data[myvars]
kc_final <- kc_new[!duplicated(kc_new[,'zipcode']),]

zipbycondition_final <- merge(zipbycondition_m,kc_final,by="zipcode")
zipbycondition_final[1] <- lapply(zipbycondition_final[1], as.factor)
```

```{r, fig.width=13,fig.height=6}
states <- map_data("state")
washington <- subset(states, region %in% c("washington"),zoom=14)

p20 <- ggplot(washington, aes(long, lat)) +
    geom_map(map=washington, aes(map_id=region), fill=NA, color="black") +
    coord_quickmap()
p20 + geom_scatterpie(aes(x=long, y=lat, group=zipcode),
    data=zipbycondition_final, cols=c("X1","X2","X3","X4","X5"), color=NA, alpha=.8) + 
    coord_fixed(xlim = c(-121.8, -122.5),  ylim = c(47.18, 47.8), ratio = 1.3) +
    labs(title="House condition distribution in the differnet zipcodes")
```


In this plot I am trying to figure how the house condition is distributed in the 
different zipcodes. The most common condition is 3 and it has the majority in
most zipcodes followed by 4 condition. 


```{r}
center_lon = median(data$long,na.rm = TRUE)
center_lat = median(data$lat,na.rm = TRUE)

facclr <- colorFactor(c("black","blue","gold","red","green"),
                      kc_data$price_cat)

m <- leaflet(kc_data) %>% addProviderTiles("Esri.NatGeoWorldMap") %>%
  addCircles(lng = ~long, lat = ~lat, 
             color = ~facclr(price_cat))  %>%
  setView(lng=center_lon, lat=center_lat,zoom = 12) %>%
  
  addLegend("bottomright", pal = facclr, values = ~price_cat,
            title = "House Price Distribution",
            opacity = .8)
m
saveWidget(m, file="map1.html")
```


This visualization predicts that the houses near the coastline are more 
expensive than the houses that are not. 


```{r}
facclr <- colorFactor(c("black","blue","gold","red","green"),
                      kc_data$yr_builtR)

m <- leaflet(kc_data) %>% addProviderTiles("Esri.NatGeoWorldMap") %>% 
  addCircles(lng = ~long, lat = ~lat, 
             radius = ~ifelse(kc_data$waterfront == 0, 5, 200),
             color = ~facclr(yr_builtR))  %>%
  setView(lng=center_lon, lat=center_lat,zoom = 15) %>%
  
  addLegend("bottomright", pal = facclr, values = ~yr_builtR,
            title = "House Built Distribution",
            opacity = .8)
saveWidget(m, file="map2.html")
```


In this plot I wanted o see how the houses were built over the years. In the 
Seattle region the houses were built near the coastline first. Then as time
progressed people started moving towards the mainland. In the right side of the
map we see most houses built after 1950s indicating newer settlemets than the
left side. Also I have addded waterfront as size so we can identify houses 
with waterfront. 

### Talk about some of the relationships you observed in this part of the \
### investigation. Were there features that strengthened each other in terms of \
### looking at your feature(s) of interest?

The costlier houses are built along the coastline and they are built in the 1900. 
The prices have remained high for those houses.

The mean prices of the houses that were built in 2000 to 2012 was way lower that 
those renovated in that period. But there is a gradual shift and the mean price 
built has increased after 2012.

### Plot One
```{r echo=FALSE, Plot_One}
ggplot(aes(x=sqft_living,y=price),data=kc_data) + geom_point(color="pink") +
  stat_smooth(method="lm",se=F,alpha=0.6,size=1) + 
  labs(title="correlation of price vs sqft living",y="Price in dollars", x="Living size in sqft")
options(scipen=999)
```

### Description One
I choose this as price has the as this gives us the strongest correlation with
sqft_living(0.71). 

### Plot Two
```{r echo=FALSE, Plot_Two, fig.width=10,fig.height=7}
spineplot(y=kc_data$price_cat,x=as.factor(kc_data$yr_builtR),
          xlab="year category", ylab="price category",main="Price Distribution over years")
```

### Description Two
The construction of houses in the cheap category has gradually reduced and 
nowerdays cheap category is not built at all. It is edging towards the expensive 
side now in the 2000s. The moderate price category was maximun in the 1951-1975 
range. 


### Plot Three
```{r echo=FALSE, Plot_Three}
facclr <- colorFactor(c("black","blue","gold","red","green"),
                      kc_data$price_cat)

m <- leaflet(kc_data) %>% addProviderTiles("Esri.NatGeoWorldMap") %>%
  addCircles(lng = ~long, lat = ~lat, 
             color = ~facclr(price_cat))  %>%
  setView(lng=center_lon, lat=center_lat,zoom = 12) %>%
  
  addLegend("bottomright", pal = facclr, values = ~price_cat,
            title = "House Price Distribution",
            opacity = .8)
m
saveWidget(m, file="map1.html")
```

### Description Three

The houses near the coastline is the cosliest and also built earlier in time. 
This indicates that people first inhabited the coastline and then in the later 
years more towards the interior in King County. 

------

# Reflection
First for doing this project comming up with a tidy dataset with lots of 
features was a challenge. I did a lot of research and found this tidy dataset on
Kaggle which suited my requirements.
The King county house pricing dataset has 21613 observations and 21 features. I 
wanted to explore all the 21 features and started by plotting the individual 
variables in the data set. Then I looked for any interesting relationships 
present in the dataset mong the features. I had expected that price was 
positively correlated with a lot of variables. But I was wrong as value for 
number of bedrooms,zipcode, lot size  was lot. The most important correlation 
is sqft of living. I found out the that condition 3 is the most prevelant among 
all houses. The houses were first built along the coastline and then people 
moved towards the mainland. 
I loved working with this housing prices data. It would have been great if the 
dataset contained day and month for the built year. I could have drawn a 
lot of insights and also use time series forcasting to predict for the next
years to come. I would continue with price prediction later on for this dataset.
I plan to do both of these things if possible with this dataset.
